---
title: "1-way ANOVA"
author: "Amnist.O"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assessment:
The file EucalyptiANOVA1.csv contains information on heights recorded for the trees of three species of Eucalyptus: E. cloeziana, E. pilularis and E. dunnii.

Perform ANOVA to test whether the average population height for at least one of the species is different from the other two.

```{r load package, include=F, echo=F}
library(tidyverse)
library(multcomp)
```

```{r load file}
df <- read.csv("datasets/EucalyptiANOVA1.csv")
df$spp <- factor(df$spp)
sample_n(df,5)
```

 - Which sample has the highest mean height?
```{r mean height}
df %>%
  reframe(across(hgt,list(n=length,mean=mean,sd=sd)),
          .by=spp)
```
 Sample with highest mean is pil with height mean of `r tapply(df$hgt,df$spp,mean)["pil"]` meters.

 - Do the residuals appear normally distributed?
```{r residual}
m1 <- lm(hgt ~ spp - 1,data = df)
plot(m1,which = 2)
```
 From the plot, the residuals is not normally distributed at the tails.

 - What is the p-value (4 decimals) of your F-test (number entry)?
 
```{r}
summary(update(m1,.~.+1))
anova(update(m1,.~.+1))
```

 Number entry means include coefficient to the linear regression. From `anova`, we get p-value of the F test at 4 decimals equals to `r anova(update(m1,.~.+1))[1,5]`