---
title: "Non-Parametric Bootstrap"
author: "Amnist.O"
date: "`r Sys.Date()`"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assessment:
Consider again the dataset EucalyptusTtest.csv, which contains information on heights (m) of trees of two species of Eucalyptus trees: E. cloeziana and E. dunnii. 

In Statistical Inference: Central Limit Theorem (Unit 2, Part 4) we used the Central Limit Theorem to produce the estimate, the standard error and the 95% confidence interval for the population mean height.

Now use non-parametric bootstrap to do the same, and compare the results.

You may also want to see how the number of bootstrapped samples affects your results.

```{r load packages, echo=FALSE}
library(tidyverse)
```

The statistical mean, se, and 95% CI from CLT are as followed:

```{r load data and CLT}
df <- read.csv("datasets/EucalyptusTtest.csv")
df %>%
  mutate(spp = factor(spp)) -> df

df %>%
  reframe(across(hgt,list(mean=mean,sd=sd,n=length)),.by = spp) %>%
            mutate(se = hgt_sd/sqrt(hgt_n),
                   lo95 = hgt_mean-(1.96*se),
                   hi95 = hgt_mean+(1.96*se)) %>%
  column_to_rownames("spp") %>%
  t() %>%
  data.frame() -> clt
clt
```

Now let's try using bootstrap to find the mean, se, and 95% CI. First, let's try a small number of bootstrap samples: 10

```{r bootstrap}
bst <- function(n) {
  bs <- data.frame()
  
  set.seed(123456)
  
  for (i in 1:1000) {
    df %>%
      reframe(hgt = sample(hgt, n,replace = T), .by = spp) -> s1
    bs %>%
      bind_rows(s1 %>%
                  reframe(hgt = mean(hgt), .by = spp)) -> bs
  }
  
  bs %>%
    reframe(across(hgt, list(
      mean = mean, sd = sd, n = length
    )), .by = spp) %>%
    mutate(
      se = hgt_sd / sqrt(hgt_n),
      lo95 = hgt_mean - (1.96 * se),
      hi95 = hgt_mean + (1.96 * se)
    ) %>%
    column_to_rownames("spp") %>%
    t() %>%
    data.frame() %>%
    rename(bs.clo = clo, bs.dun = dun) %>%
    bind_cols(clt) %>%
    mutate(across(everything(),  ~ formatC(
      .x, digits = 6, format = "f"
    )))
}
bst(10)
```

Now let's try a sample same as original $n=60$
```{r}
bst(60)
```

Finally, a very large sample of $n=1000$
```{r}
bst(1000)
```

Higher sample size leads to more accurate mean and variance