---
title: "Central Limit Theorem"
author: "Amnist.O"
date: "`r Sys.Date()`"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assessment: 
The dataset EucalyptusTtest.csv contains information on heights (m) of trees of two species of Eucalyptus trees: E. cloeziana and E. dunnii.

Do some exploratory analysis:

 - Evaluate sample statistics for each species.
 - Produce a histogram/density plot for each species.
Use the Central Limit Theorem to produce
 - the estimate
 - the standard error
 - the 95% confidence interval for the population mean height.

```{r load data, echo=FALSE}
library(tidyverse)
df <- read.csv("datasets/EucalyptusTtest.csv")
```

```{r exploratory}
df %>%
  mutate(spp = factor(spp)) -> df

summary(df)

df %>%
  reframe(across(hgt,list(mean=mean,sd=sd,n=length)),.by = spp) %>%
            mutate(se = hgt_sd/sqrt(hgt_n),
                   lo95 = hgt_mean-(1.96*se),
                   hi95 = hgt_mean+(1.96*se)) %>%
  column_to_rownames("spp") %>%
  t() %>%
  data.frame()
```

Consider a seed germination experiment, in which 15 out of 20 seeds have germinated.

Use CLT to produce:

 - The estimate for the population proportion
 - The associated standard error
 - The 95% confidence interval
 
```{r seed germination}
seed <- c(rep(1,15),rep(0,5))

cat("sample size =",length(seed),"\n")
cat("sd = ",sd(seed),"\n")
cat("mean = ",mean(seed),"\n")
cat("se = ",sd(seed)/sqrt(length(seed)),"\n")
cat("lo95 = ",mean(seed)-(1.96*sd(seed)/sqrt(length(seed))),"\n")
cat("hi95 = ",mean(seed)+(1.96*sd(seed)/sqrt(length(seed))),"\n")
```
 
