---
title: "t-test"
author: "Amnist.O"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assessment:
The dataset EucalyptusTtest.csv contains information on heights (m) of trees of two species of Eucalyptus trees: E. cloeziana and E. dunnii.

Use the t-test in R to test the hypothesis that: there is no difference in average height for the two species.

You may find the following steps useful:

 - Evaluate species-specific sample statistics and produce a boxplot. What do you see?
 - Check the normality of samples. Does it matter in this situation? Why or why not?
 - Perform the t-test and interpret the results.
 - Try logging the response and repeating the above steps.
 
```{r import data and packages, include = F, echo=FALSE}
library(tidyverse)
df <- read.csv("datasets/EucalyptusTtest.csv")
```

## Sample statistics and boxplot

```{r}
df$spp <- factor(df$spp)
df %>%
  ggplot(aes(x=spp,y=hgt,fill=spp))+
  geom_boxplot()

df %>%
  reframe(across(hgt,list(n=length,mean=mean,sd=sd)),.by = spp) %>%
            mutate(se = hgt_sd/sqrt(hgt_n),
                   lo95 = hgt_mean-(1.96*se),
                   hi95 = hgt_mean+(1.96*se)) %>%
  column_to_rownames("spp") %>%
  t() %>%
  data.frame()
```

## Sample normality

```{r qq plot}
qqnorm(df$hgt[df$spp=="clo"],main="clo")
qqline(df$hgt[df$spp=="clo"],col="turquoise3")
qqnorm(df$hgt[df$spp=="dun"],main="dun")
qqline(df$hgt[df$spp=="dun"],col="turquoise3")
```

## t-test of the sample

```{r t-test}
t.test(hgt~spp,data=df)
```

t-test shows that p-value = `r t.test(hgt~spp,data=df)$p.value` meaning that the difference in both sample means is not statistically significant. The 95% CI is `r t.test(hgt~spp,data=df)$conf.int[1:2]`.

## Log-transformed of the sample

Statistics and boxplot of log transformed data are as followed:

```{r}
df %>%
  mutate(hgt = log(hgt)) -> ldf

ldf %>%
  reframe(across(hgt,list(n=length,mean=mean,sd=sd)),.by = spp) %>%
            mutate(se = hgt_sd/sqrt(hgt_n),
                   lo95 = hgt_mean-(1.96*se),
                   hi95 = hgt_mean+(1.96*se)) %>%
  column_to_rownames("spp") %>%
  t() %>%
  data.frame()

ldf %>%
  ggplot(aes(x=spp,y=hgt,fill=spp))+
  geom_boxplot()
```

Normality of the log transformed data are displayed with qqnorm as followed:
```{r}
qqnorm(ldf$hgt[ldf$spp=="clo"],main="clo")
qqline(ldf$hgt[ldf$spp=="clo"],col="lawngreen")
qqnorm(ldf$hgt[ldf$spp=="dun"],main="dun")
qqline(ldf$hgt[ldf$spp=="dun"],col="springgreen3")
```

Log transformed does not help that much.

t-test of the log transformed data are as followed:
```{r}
t.test(hgt~spp,data=ldf)
```

Interpreting log transformed data need some extra calculation since the value are in log form. We just need to exponent them with natural number $e$ and their difference are percentage since it is $e^{log(a)-log(b)} = a/b$
