---
title: "Multinomial logit assessment"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assessment: Nominal categorical GLM
Repeat the analysis in this section (first, perform the 1-way ANOVA, then fit a multinomial logistic model) using the variable `crownarea` rather than the `hgt`.

```{r load packages and data, message=FALSE}
library(tidyverse)
library(nnet)

d <- read.csv(file = "datasets/eucalyptus.csv")
anova(lm(crownarea ~ spp, data = d))
```

```{r model}
m <- multinom(spp ~ crownarea, data = d)
summary(m)
```

To test the coefficients, we model without the coefficient and compare both models with anova

```{r compare coef}
m0 <- multinom(spp ~ 1, data = d)
anova(m0,m)
lapply(list("no coef" = m0, coef = m), AIC)
```

It appears that there is a statistically significant difference between the two models. The crownarea of the tree does affect the probability of belonging to a particular species.

## Plots
Let's plot the result of multinomial:

```{r plot}
d.pred <- data.frame(crownarea = seq(8,200,1))

matrix.prob <- predict(m,d.pred, type ="probs")

d2 <- expand.grid(crownarea = seq(8,200,1),
                  Species = c("clo","dun","pil"))

d2$p <- c(matrix.prob)

d2 %>%
  ggplot(aes(x = crownarea, y = p, group = Species))+
  geom_line(aes(col = Species),linewidth = 1.5)+
  xlab("Crownarea")+
  ylim(c(0,1))
```

Based on the graph, trees should be `dun` when the crownarea is more than around 50, `pil` when the crownarea is between 8 - 50, and `clo` otherwise.

## Goodness of fit
We can check the goodness of fit of the model with apparent error ratio (AER). Let's start by recording the predicted values and creating the *confusionn matrix*.

```{r pred vs obs}
pred <- predict(m)
table(d$spp,pred)
```

The confusion matrix is a contingency table of observed vs. predicted data. The diagonal is the correctly predicted cases. Others are errors. So, we can get AER by $AER = \frac{error}{total}$.

```{r aer}
scales::percent(mean(pred != d$spp),.01)
```

## Alternative models
Now try using a transformation of the crownarea, such as $\log()$ or a square root.

Try a model with both, crownarea and hgt, and their interaction.

Is your AER for that model better than for the crownarea-only one?

```{r alternative models}
m.int <- multinom(spp ~ crownarea*hgt, data = d)
m.add <- multinom(spp ~ crownarea+hgt, data = d)
m.log <- update(m,.~log(.))
m.sqrt <- update(m,.~sqrt(.))

aer <- function(m, df) {
  pred <- predict(m)
  table(df$spp, pred)
  result <- mean(pred != df$spp)
  return(result)
}

lapply(list(base = m,
            log = m.log,
            sqrt = m.sqrt,
            add = m.add,
            int = m.int), aer,d)
```

