---
title: "Ordered logit assessment"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assessment
Use the lego dataset to test whether the effect of age on the assigned strength of emotion is modified by the sex of the participant.

```{r load packages and data, message=FALSE}
library(tidyverse)
library(MASS)

d <- read.csv("datasets/lego.csv")
d$strength <- factor(d$strength,ordered = T)
summary(d)
```

```{r model}
m.add <- polr(strength ~ age+sex, data = d,subset = !is.na(age))
m.int <- update(m.add,.~age*sex)
anova(m.add,m.int)
```

In other words are the models strength ~ age*sex and strength ~ age+sex statitsically significantly different?  
 - From the anova, the models are not statistically significantly different.


 Use the model output to produce the plot of the cumulative probabilities for each strength category by age and sex.

```{r plot}
expit <- function(x){1/(1+exp(-(x)))}

par(mfrow=c(1,2))
# male
curve(expit(m.int$zeta[1]+coef(m.int)[1]*x+coef(m.int)[2]+coef(m.int)[3]*x),
      from=0,to=100,col='purple',
      ylim=c(0,1),xlab='Age',ylab='Pr', main = "male")
curve(expit(m.int$zeta[2]+coef(m.int)[1]*x+coef(m.int)[2]+coef(m.int)[3]*x),
      from=0,to=100,col='cornflowerblue',add=T)
curve(expit(m.int$zeta[3]+coef(m.int)[1]*x+coef(m.int)[2]+coef(m.int)[3]*x),
      from=0,to=100,col='lawngreen',add=T)
curve(expit(m.int$zeta[4]+coef(m.int)[1]*x+coef(m.int)[2]+coef(m.int)[3]*x),
      from=0,to=100,col='orange',add=T)
abline(h=0, col='sienna'); abline(h=1, col='sienna')

# female
curve(expit(m.int$zeta[1]+coef(m.int)[1]*x),
      from=0,to=100,col='purple',
      ylim=c(0,1),xlab='Age',ylab='Pr', main = "female")
curve(expit(m.int$zeta[2]+coef(m.int)[1]*x),
      from=0,to=100,col='cornflowerblue',add=T)
curve(expit(m.int$zeta[3]+coef(m.int)[1]*x),
      from=0,to=100,col='lawngreen',add=T)
curve(expit(m.int$zeta[4]+coef(m.int)[1]*x),
      from=0,to=100,col='orange',add=T)
abline(h=0, col='sienna'); abline(h=1, col='sienna')

```

Produce the confusion matrix. What do you think about the fit of your model? Is it a good/useful model?

```{r confusion matrix}
pred <- predict(m.int)
table(m.int$model$strength,pred, dnn = list("Actual","Predicted"))
mean(as.numeric(pred) != as.numeric(m.int$model$strength))
```

The model is pretty bad. It cannot predict other strength other than 2. The error rate is 73.38%.

Based on your modeling, what is the probability that a 40 year-old man will assign an emotion of random lego figurine into the maximum strength (5) category?

```{r prob}
predict(m.int,newdata=data.frame(age = 40,sex = "M"), type = "probs")
```

According to the model, the probability of a 40 year-old man assigning an emotion of random lego figurine into emotion strength of 5 is `r scales::percent(predict(m.int,newdata=data.frame(age=40,sex="M"),type="probs")[5],.01)`.