---
title: "Probit link function"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "D:/git/ucx-stat/probit_files/figure-gfm/")
```

## Assessment
There is another link function: probit. Probit is an inverse of the standard cumulative normal density function:

$$
\Phi(p) = \int_0^p \frac{1}{\sqrt{2\pi}}e^{-\frac{1}{2}x^2}dx
$$
We can plot using `curve(qnorm(x),from = 0, to = 1, lwd = 2)`
```{r probit plot}
curve(qnorm(x),from = 0, to = 1, lwd = 2)
```

while logit looks like:
```{r plot logit}
curve(log(x/(1-x)), from = 0, to = 1, lwd = 2, ylab = "logit")
```

plotting both together makes:
```{r plot together}
curve(qnorm(x), from = 0, to = 1, lwd = 2, col = "salmon",ylab = "y")
curve(log(x/(1-x)), from = 0, to = 1, lwd = 2, col = "blue", add = T)
legend(x = 0.02, y = 2, legend = c("probit","logit"), col = c("salmon","blue"), lwd = 2)
title("probit vs logit curves")
```

## Repeating the ANCOVA

Repeat the ANCOVA analysis to test whether the probability of winning depends on the Chaser after adjusting for the team score. Use probit instead of logit as a link function.
```{r load packages and data, message=FALSE}
library(tidyverse)
chase <- read_lines(file = "datasets/chase")
df <- read.table(file = chase, header = T, sep = ",", stringsAsFactors = T)
```

```{r model}
p <- glm(Win ~ TeamScore * Chaser, data = df,
         family = "binomial"(link = "probit"))
m <- update(p, family = "binomial"(link = "logit"))
modelsummary::msummary(list(logit = m,probit = p), stars = T)

# AER logit
cat("AER logit\n")
table(df$Win,predict(m,type = "response")>.5,dnn = list("Actual","Predicted"))


#AER probit
cat("\nAER probit\n")
table(df$Win,predict(p,type = "response")>.5,dnn = list("Actual","Predicted"))
```

Once you have completed your analysis, please answer the following questions:
Are the ANCOVA results qualitatively different?
```{r ancova}
anova(m,update(m,.~TeamScore),test="Chisq")
anova(p,update(p,.~TeamScore),test="Chisq")
```

 - We can see from the ancova that the probability of winning does not depend on chaser after adjusting team score in both logit model and probit model.

What method can be used to compare the two models with different link functions?
 - Since the response is the same, we can use AIC to compare between models. Also, AER can also be used to compare since it measures the correctness of the models.

Which model is statistically better?
 - From the AIC (the lesser, the better), we can state that probit model is better ($`r AIC(p)`<`r AIC(m)`)$.

What is the AIC for the GLM with the probit link function?
 - AIC of the probit link function is `r AIC(p)` for additive model and `r AIC(update(p,.~TeamScore*Chaser))` for interaction model (question refers to interaction model).

Plot the curves for chaser Anne based on both models
```{r curve plot}
curve(1/(1+exp(-(coef(m)[1]+coef(m)[2]*x))),from = 0, to = 30,lwd = 2, col = "salmon",
      ylab = "Winning probability")
lines(x = seq(0,30,0.5),
      y = predict(p,list(TeamScore = seq(0,30,0.5),Chaser = rep("Anne",61)),type = "response"),
      col = "blue", lwd = 2, lty = 2)
title("logit vs probit winning probability against chaser Anne")
legend(x = 0.02, y= 0.975,
       legend = c("logit","probit"),
       col = c("blue","salmon"),
       lwd = 2,
       lty = c(2,1))
```

 - They look almost the same.