---
title: "Repeated Measures ANOVA"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assessment
 - Load the eucalyptus.csv data and repeat the analysis of the previous section with the diameter at breast height `dbh` rather than height `hgt` as the response variable.

```{r load packages and data, message=FALSE}
library(tidyverse)
library(lme4)
library(lmerTest)
# load data
df <- read.csv("datasets/eucalyptus.csv")
df$stocking <- factor(df$stocking)
```

```{r model adjust for species}
m <- lmer(dbh ~ spp*stocking + (1|plot), data = df)
m1 <- lmer(dbh ~ stocking + (1|plot), data = df)
anova(m,m1)
```

The anova test shows that the effect of stocking is statistically significant on dbh after adjusting for species.

```{r adjsut for stocking}
s <- lmer(dbh ~ spp+(1|plot),data = df)
anova(m,s)
```

The anova test shows that the effect of species is statistically significant on dbh after adjusting for stocking.

```{r stocking x species}
m.add <- update(m,.~spp+stocking+(1|plot))
anova(m,m.add)
```

The anova test shows that the effect of stocking differed by species. 