---
title: "Linear regression with mixed type of variables"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Linear regression with one continuous and one binary input
Let's try to combine what we have done by inspecting the association between weight and height with sex.

```{r load package, message=FALSE}
library(tidyverse)
```

```{r laod data and set up model}
df <- read.csv("datasets/howell1.csv",sep=";")
df %>%
  filter(age >=18) -> df
```

### Additive model
Additive model is the simplest model:
$$\mathbb{E}(weight|height) = \beta_0+\beta_1height +\beta_2sex$$
$\beta_1$ is the average change in weight associated with a 1 cm increase in height and $\beta_2$ is the difference in weight between a man and a woman of the same height in every height.

For female the model looks like:
$$\mathbb{E}(weight|height)=\beta_0+\beta_1height$$

For male, the model looks like:
$$\mathbb{E}(weight|height)=\beta_0+\beta_2+\beta_1height$$

In graphical representation, the line for male and female will be $\beta_2$ apart regardless of height level.

Let's fit the model
```{r additive model}
addm <- lm(weight ~ height+male,data=df)
summary(addm)
```

Let's plot the predicted lines and the 95% CI envelopes.
```{r fitted plot}
visreg::visreg(addm,
               xvar = "height",
               by="male",
               overlay=T)
```

or calculate by hands:
```{r}
#preparing a grid for predicted values:
# a range of heights for both sexes:
d.pred <- expand.grid(height = seq(130, 180, 1), 
                      male = c(0, 1))

# obtaining predictions based on the m1
d.pred$weight <- predict(addm, newdata = d.pred)

# obtaining the standard errors for those predictions:
d.pred$weight.se <- predict(addm, newdata = d.pred, 
                            se.fit = T)$se

library(ggplot2)
ggplot(data = df, aes(x = height, y = weight)) +
  geom_point(aes(group = male, col = factor(male))) +
  geom_ribbon(
    data = d.pred,
    aes(
      x = height,
      ymin = weight - 1.96 * weight.se,
      ymax = weight + 1.96 * weight.se,
      group = male,
      fill = factor(male)
    ),
    alpha = .3
  ) +
  geom_line(data = d.pred,
            aes(
              x = height,
              y = weight,
              group = male,
              col = factor(male)
            ),
            linewidth = 1.2)
```

```{r}
summary(addm)
```

From the result, we can see that the fitted line for male is below female in every height indicated that male has less weight than female of the same height.

## Linear regression with interaction
Additive model assumed that the association between height and weight is the same for men and women. However, the association of height and weight change according to sex.

$$\mathbb{E}(weight|height)=\beta_0+\beta_1height+\beta_2sex+\beta_3 \space height\times sex$$

The term $\beta_3\space height \times sex$ is called interaction term. When $sex=0$, $\beta_3$ has no effect to the weight.

```{r}
intm <- lm(weight ~ height+male+height:male,data=df)
visreg::visreg(intm,xvar="height",by="male",overlay=T)
```

calculation by hand the plot become:
```{r}
# obtaining predictions based on the m2
d.pred$weight2 <- predict(intm, newdata = d.pred)

# obtaining the standard errors for those predictions:
d.pred$weight.se2 <- predict(intm, newdata = d.pred, se.fit = T)$se

ggplot(data = df, aes(x = height, y = weight)) +
  geom_point(aes(group = male, col = factor(male))) +
  geom_ribbon(
    data = d.pred,
    aes(
      x = height,
      ymin = weight2 - 1.96 * weight.se2,
      ymax = weight2 + 1.96 * weight.se2,
      group = male,
      fill = factor(male)
    ),
    alpha = .3
  ) +
  geom_line(data = d.pred,
            aes(
              x = height,
              y = weight2,
              group = male,
              col = factor(male)
            ),
            linewidth = 1.2)
```
```{r}
summary(intm)
```

We can see that the slopes for males is a bit smaller than that of female. The difference between slope is $\beta_3 = -0.04243$. However, the difference is not significant (p value > 0.05)

## Common model vs Individual models
First, regression model assumes that the measurement error $\varepsilon_i$ are normally distributed with common variance. With individual models, we allow this variance to be different for men and women.

Secondly, separating the dataset does not allow comparison betwenn slope coefficients. Whereas fitting common model allows us to say that the effect of height on weight was on average 42g/cm smaller for men than for women (p=0.608)

Remember to check the assumptions for the analysis.