---
title: "Multinomial logit"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let's fit the multinomial logit model on the eucalyptus data.

```{r load packages and data, message=FALSE}
library(nnet)
library(tidyverse)

d <- read.csv("datasets/eucalyptus.csv")
anova(lm(hgt ~ spp, data = d))
```

ANOVA found that the differences between species in terms of average height is statistically significant. So, let's use height to classify the trees into species with multinomial logit.

```{r model}
m <- multinom(spp ~ hgt, data = d)
summary(m)
```

There are the coefficient estimates for `dun` and `pil` but none for `clo` since `clo` is the reference. There are no p values in the table as well. This is because it makes no sense to test individual coefficients. If the variable $x$ is in the model, we have to *simultaneously* test $G-1$ coefficients where $G$ is the total number of groups.

To test the coefficients, we model without the coefficient and compare both models with anova

```{r compare coef}
m0 <- multinom(spp ~ 1, data = d)
anova(m0,m)
lapply(list("no coef" = m0, coef = m), AIC)
```

It appears that there is a statistically significant difference between the two models. The height of the tree does affect the probability of belonging to a particular species.

## Plots
Let's plot the result of multinomial:

```{r plot}
d.pred <- data.frame(hgt = seq(10,25,.1))

matrix.prob <- predict(m,d.pred, type ="probs")

d2 <- expand.grid(hgt = seq(10,25,.1),
                  Species = c("clo","dun","pil"))

d2$p <- c(matrix.prob)

d2 %>%
  ggplot(aes(x = hgt, y = p, group = Species))+
  geom_line(aes(col = Species),linewidth = 1.5)+
  xlab("Height (m)")+
  ylim(c(0,1))
```

Based on the graph, trees should be `dun` when the height is less than around 15 m., `pil` when the height is between 15 - 16 m., and `clo` otherwise.

We can check the goodness of fit of the model with apparent error ratio (AER). Let's start by recording the predicted values and creating the *confusionn matrix*.

```{r pred vs obs}
pred <- predict(m)
table(d$spp,pred)
```

The confusion matrix is a contingency table of observed vs. predicted data. The diagonal is the correctly predicted cases. Others are errors. So, we can get AER by $AER = \frac{error}{total}$.

```{r aer}
mean(pred != d$spp)
```

So, $`r scales::percent(mean(pred != d[,"spp"]),.01)`$ of all observations is incorrectly predicted. Not so good model...

Even the 1-way anova tells us that the differences between species and height are highly statistically significant on the probability of belonging to a particular species, the classification results still does not perform well.

**Statistical significance is not everything.**