---
title: "Ordered Logit"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Fitting proportional hazards ordinal logit regression
We use data on people's perception of emotion of Lego faces.

```{r load packages and data, message=FALSE}
library(tidyverse)
library(MASS)

d <- read.csv("datasets/lego.csv")
# define strength of emotion as ordered factor
d$strength <- factor(d$strength, ordered = T)
summary(d)
```

## Fitting the model
```{r model}
m <- polr(strength ~ age, data = d)
summary(m)
```

From the proportion hazards, the slope of the independent variable is equal: $\beta_1 = \beta_2 = \ldots =\beta_n$ where $n$ is number of ordinal rank - 1.

```{r plot}
expit <- function(x){1/(1+exp(-(x)))}

curve(expit(-1.6312+0.0017*x),from=0,to=100,col='purple',
      ylim=c(0,1),xlab='Age',ylab='Pr')
curve(expit(-0.2528+0.0017*x),from=0,to=100,col='cornflowerblue',add=T)
curve(expit( 0.7301+0.0017*x),from=0,to=100,col='lawngreen',add=T)
curve(expit( 1.8279+0.0017*x),from=0,to=100,col='orange',add=T)
abline(h=0, col='sienna'); abline(h=1, col='sienna')
```

Seems like age does not matter in this case.

## Interpretation of the coefficients
The coefficient of the age is `r coef(m)`. How do we interpret this. Recall the log-odds interpretation:  
*Each additional year of age is associated with an average 0.17% increase in the odds of choosing a higher strength category.*

Is the effect statistically significant? The model does not give us p-value, but we can use `anova()` for the model with and without the variable of interest.

```{r anova}
anova(m,update(m,.~1,data = d, subset= !is.na(age)))
```

ANOVA tells us that age does not have a significant effect on percieved strength of emotion.

## More complex model
We can use other factor as well. Let's see the association between age and emotion to strength of emotion perceived:

```{r int model}
m2 <- polr(strength ~ age*emotion, data = d)
m3 <- update(m2,.~emotion,subset=!is.na(age))
anova(m2,m3)
```


